Chemical tagging can work: \\ Identification of stellar phase-space structures \\ purely by chemical-abundance similarity\vspace{-7ex}

### 2016-08-25 papers/kmeans/kmeans.tex
pull back on density

When the k-means results are returned, the $D\times D$ empirical
variance tensor for the members of each cluster can be constructed.
From this, an effective density in the abundance space can be computed
as the number of points in the cluster divided by the square-root of
the determinant of the tensor.
This density was used to rank abundance-space overdensities (from highest
to lowest) for visual inspection.
In \figurename~\ref{fig:densities} we show the distribution of cluster membership
and density for the $K=256$ run of k-means.

@@

\begin{figure}[!bp]
\includegraphics[width=1.5\figwidth]{./figs/clusters_0256.pdf}
\caption{The distribution of membership and density for the 256
  abundance-space clusters returned by the k-means algorithm at
  $K=256$.  The densest cluster (which is displayed in more detail in 
  \figurename~\ref{fig:M13}) is far more dense than the second
  densest.\label{fig:densities}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0253}
\caption{Same as \figurename~\ref{fig:all} except that the entire
  sample has been made gray and the members of the most dense $K=256$
  abundance-space cluster have been rendered as unique, prominent
  triangles (color is velocity rank, orientation is $\logg$ rank).
  The lower-right plot shows this cluster (circle) in context of the
  other clusters (dots).  This cluster, which was identified only in
  abundance space (six projections of which are the top six panels in
  this \figurename), turns out to be dominated by the halo globular
  cluster M13. The symbol orientations and colors have no meaning;
  they are randomly generated---one unique color and orientation for
  each highlighted star---to make the points cross-identifiable across
  panels.\label{fig:M13}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0034}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by globular
  cluster M5.\label{fig:M5}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0177}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by the Sagittarius dwarf spheroidal galaxy.\label{fig:Sgr}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0010}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by a hitherto unrecognized high-velocity-dispersion
  structure in the Galaxy halo.\label{fig:halo}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0141}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by a very thin young stellar
  structure in the Galaxy disk, probably one that has been discovered previously \citep{wegg}.\label{fig:disk}}
\end{figure}

@@

It is conceivable that this structure, being near the outskirts of the
abundance distribution, is somehow concentrated in abundance space by
some kind of shrinkage (in the statistical sense) induced by
\thecannon's regression.
However, it is a very dense feature in abundance space, and is worth
following up.
If follow-up observations show that the stars have abundances and that
are consistent with having origin in a single stellar population or
low-metallicity dwarf galaxy, it could represent an accretion event in
the Milky Way halo.

>>>>

When the k-means results are returned, the $D\times D$ empirical
variance tensor for the members of each cluster can be constructed.
From this, an effective density in the abundance space can be computed
as the number of points in the cluster divided by the square-root of
the determinant of the tensor.
This density was used to rank abundance-space overdensities
for visual inspection.
In \figurename~\ref{fig:densities} we show the distribution of cluster membership
and density for the $K=256$ run of k-means.
There is a bulk trend of larger clusters (clusters with more members or higher occupation number)
being more dense, but the most interesting
k-means densities are those that are more dense than this trend.

@@

\begin{figure}[!bp]
\includegraphics[width=1.5\figwidth]{./figs/clusters_0256.pdf}
\caption{The distribution of membership and density for the 256
  abundance-space clusters returned by the k-means algorithm at
  $K=256$.  There is a bulk trend and then clusters that are much more
  dense than the trend. The densest cluster is displayed in more detail in 
  \figurename~\ref{fig:M13}.\label{fig:densities}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0253}
\caption{Same as \figurename~\ref{fig:all} except that the entire
  sample has been made gray and the members of the most dense $K=256$
  abundance-space cluster have been rendered as unique, prominent
  triangles (color is velocity rank, orientation is $\logg$ rank).
  The lower-right plot shows this cluster (circle) in context of the
  other clusters (dots).  This cluster, which was identified only in
  abundance space (six projections of which are the top six panels in
  this \figurename), turns out to be dominated by the halo globular
  cluster M13. The symbol orientations and colors have no meaning;
  they are randomly generated---one unique color and orientation for
  each highlighted star---to make the points cross-identifiable across
  panels.\label{fig:M13}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0034}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by globular
  cluster M5.\label{fig:M5}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0177}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by the Sagittarius dwarf spheroidal galaxy.\label{fig:Sgr}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0010}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by a hitherto unrecognized high-velocity-dispersion
  structure in the Galaxy halo.\label{fig:halo}}
\end{figure}
\begin{figure}[!p]
\insanefigure{cluster_0256_0141}
\caption{Same as \figurename~\ref{fig:M13} but for another dense
  abundance-space cluster.
  This one turns out to be dominated by a very thin young stellar
  structure in the Galaxy disk, probably one that has been discovered previously \citep{wegg}.\label{fig:disk}}
\end{figure}

@@

It is conceivable that this structure, being near the outskirts of the
abundance distribution, is somehow concentrated in abundance space by
some kind of shrinkage (in the statistical sense) induced by
\thecannon's regression.
However, it is a dense feature in abundance space, and is worth
following up.
If follow-up observations show that the stars have abundances and that
are consistent with having origin in a single stellar population or
low-metallicity dwarf galaxy, it could represent an accretion event in
the Milky Way halo.
